---
title: "Vernal Pool Analysis Markdown"
output: html_document
date: "2024-03-11"
---

```{r setup, echo=FALSE,warning=FALSE,include=FALSE}
library(ggplot2)
#First upload rain data and get max and total
noaa<-read.csv("~/VP_hydrology/Monthly_NOAA.csv")
noaayrly<-aggregate(cbind(Precip_mm,AVG_Temp_C)~Rain.Year,noaa,FUN=max)
noaayrly<-noaayrly[noaayrly$Rain.Year>=2019,]
noaayrly$wtr_yr<-c("WY19","WY20","WY21","WY22","WY23","WY24")

#Load clean vp data
vp<-read.csv("~/VP_hydrology/vp_clean.csv")
vpq<-vp[vp$Water_Level!=0,]
vpq$days<-7
vpqag<-aggregate(days~Location+VP_Name+wtr_yr,vpq,FUN=sum)
vpmax<-aggregate(Water_Level~Location+VP_Name+wtr_yr,vp,FUN=max)

vpall<-merge(vpqag,noaayrly,by="wtr_yr",all.x=TRUE,all.y=TRUE)
vpall<-vpall[!is.na(vpall$VP_Name),]
vpall$Precip_in<-vpall$Precip_mm/25.4
vpall$Precip_cm<-vpall$Precip_mm/10
lubridate::ymd(vpall$Rain.Year, truncated = 2L)
```

## Number of days inundated for each vernal pool
Each site should be sampled once a week therefore each depth measurement here is interpreteed as 7 days. All 0 entries are counted as 7 days with no inundation. The blue line represents yearly rainfall (cm).

```{r,echo=FALSE,warning=FALSE}
for (i in unique(vpall$Location)){
  p<-vpall[vpall$Location==i,]
  q<- ggplot(p,aes(x=Rain.Year)) +
    geom_bar(stat="identity", position=position_dodge(),aes(y=days, fill=VP_Name))+ylab("Total Days Inundated")+xlab("Vernal Pool Name")+ 
    theme_bw()+geom_line(color="lightblue",size=1.5,aes(y=Precip_cm))+geom_point(aes(y=Precip_cm))+
    theme(axis.title.y = element_text(size=18,margin = margin(t = 0, r = 10, b = 0, l = 20)), 
          axis.title.y.right = element_text(size=15,margin = margin(t = 0, r = 0, b = 0, l = 20)),
          axis.title.x=element_text(size=18), axis.text.x=element_text(size=18),
          axis.text.y = element_text(size=18),title =element_text(size=18, face='bold'),
          strip.text.y = element_text(size = 18),
          legend.text = element_text(size=12),
          legend.title = element_text(size=12),
          legend.position = "right")+ggtitle(p$Location)+
    guides(fill=guide_legend(ncol=2))
  print(q)
}
```

## Maximum water depth at each vernal pool


```{r,warning=FALSE,echo=FALSE}
vpmax<-vpmax[vpmax$Water_Level<90,]
for (i in unique(vpmax$Location)){
  p<-vpmax[vpmax$Location==i,]
  q<- ggplot(p,aes(x=wtr_yr)) +
    geom_bar(stat="identity", position=position_dodge(),aes(y=Water_Level, fill=VP_Name))+ylab("Maximum Water Level")+xlab("Vernal Pool Name")+ 
    theme_bw()+
    theme(axis.title.y = element_text(size=18,margin = margin(t = 0, r = 10, b = 0, l = 20)), 
          axis.title.y.right = element_text(size=15,margin = margin(t = 0, r = 0, b = 0, l = 20)),
          axis.title.x=element_text(size=18), axis.text.x=element_text(size=18),
          axis.text.y = element_text(size=18),title =element_text(size=18, face='bold'),
          strip.text.y = element_text(size = 18),
          legend.text = element_text(size=12),
          legend.title = element_text(size=12),
          legend.position = "right")+ggtitle(p$Location)+
    guides(fill=guide_legend(ncol=2))
  print(q)
}
```

## Water Quality

```{r,echo=FALSE,Warning=FALSE,include=FALSE}
WQclean<-read.csv("~/VP_hydrology/VP_WQ_Clean.csv")
WQclean$Date<-as.POSIXct(WQclean$Date,format<-"%Y-%m-%d %H:%M:%S")
WQclean<-WQclean[WQclean$Salinity_ppt<100&WQclean$DO_mg_L<300&WQclean$Temp_C<100,]
WQclean<-WQclean[!is.na(WQclean$Salinity_ppt),]
```

## Salinity analysis

```{r,echo=FALSE,warning=FALSE}
for (i in unique(WQclean$Location)){
  p<-WQclean[WQclean$Location==i,]
  q<- ggplot(p,aes(x=Date)) +
    geom_line(size=1.5,aes(y=Salinity_ppt, color=VP_Name))+geom_point(aes(y=Salinity_ppt))+
    ylab("Salinity (ppt)")+xlab("Date")+theme_bw()+ggtitle(p$Location)+
    theme(axis.title.y = element_text(size=18,margin = margin(t = 0, r = 10, b = 0, l = 20)), 
          axis.title.y.right = element_text(size=15,margin = margin(t = 0, r = 0, b = 0, l = 20)),
          axis.title.x=element_text(size=18), axis.text.x=element_text(size=18),
          axis.text.y = element_text(size=18),title =element_text(size=18, face='bold'),
          strip.text.y = element_text(size = 18),
          legend.text = element_text(size=12),
          legend.title = element_text(size=12),
          legend.position = "right")+ggtitle(p$Location)
  print(q)
}
```

## Dissolved Oxygen Analysis

```{r,echo=FALSE,warning=FALSE}
for (i in unique(WQclean$Location)){
  p<-WQclean[WQclean$Location==i,]
  q<- ggplot(p,aes(x=Date)) +
    geom_line(size=1.5,aes(y=DO_mg_L, color=VP_Name))+geom_point(aes(y=DO_mg_L))+
    ylab("Dissolved Oxygen (mg/L)")+xlab("Date")+theme_bw()+ggtitle(p$Location)+
    theme(axis.title.y = element_text(size=18,margin = margin(t = 0, r = 10, b = 0, l = 20)), 
          axis.title.y.right = element_text(size=15,margin = margin(t = 0, r = 0, b = 0, l = 20)),
          axis.title.x=element_text(size=18), axis.text.x=element_text(size=18),
          axis.text.y = element_text(size=18),title =element_text(size=18, face='bold'),
          strip.text.y = element_text(size = 18),
          legend.text = element_text(size=12),
          legend.title = element_text(size=12),
          legend.position = "right")+ggtitle(p$Location)
  print(q)
}
```

## Temperature analysis

```{r,echo=FALSE,warning=FALSE}
for (i in unique(WQclean$Location)){
  p<-WQclean[WQclean$Location==i,]
  q<- ggplot(p,aes(x=Date)) +
    geom_line(size=1.5,aes(y=Temp_C, color=VP_Name))+geom_point(aes(y=Temp_C))+
    ylab("Temperature (C)")+xlab("Date")+theme_bw()+ggtitle(p$Location)+
    theme(axis.title.y = element_text(size=18,margin = margin(t = 0, r = 10, b = 0, l = 20)), 
          axis.title.y.right = element_text(size=15,margin = margin(t = 0, r = 0, b = 0, l = 20)),
          axis.title.x=element_text(size=18), axis.text.x=element_text(size=18),
          axis.text.y = element_text(size=18),title =element_text(size=18, face='bold'),
          strip.text.y = element_text(size = 18),
          legend.text = element_text(size=12),
          legend.title = element_text(size=12),
          legend.position = "right")+ggtitle(p$Location)
  print(q)
}


```



